Continuous Deductive Reasoning over Streamed RDF Data

### 2015-11-17 article.tex
fixed left/mid/right sizing of braces and "givens"

We define the concept of entailment in Datalog as follows. A fact $e$
is entailed by a program $\Pi$ iff each interpretation satisfying
$\Pi$ (satisfying every clause in $\Pi$) also satisfies $e$, written
$\Pi \models e$. For a program $\Pi$ and a set of entailed facts
$E$, $\Pi \models E$ iff $\Pi \models e$ for every $e \in E$. For
convenience, we refer to the set of all facts that are entailed by a
program (the intersection of all the Herbrand models of the program)
as the {\em entailment of a program}, written $cons(\Pi)$ and defined
as:
\begin{align*}
  cons(\Pi) &= \{ e \in \hat{H} \mid \Pi \models e \} \\
  &= \bigcap_\mathcal{I} \models_\mathcal{I} \Pi
\end{align*}
\end{definition}

@@

For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
of fact instances and $F = \{ strip(f) \mid f \in F^T \}$.

@@

The {\em entailment of the TA-Datalog program} $\Pi^T$ is therefore:
\[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
\Pi^T \models \langle e, t_e \rangle \} \]

@@

For an entailed fact instance $\langle e, t_e\rangle$ and a
justification of that fact instance $J^T$, the instant $t_e$ from which
$\langle e, t_e \rangle$ is recognised to be true is, intuitively, taken to be the
latest instant from which any of the fact instances in the justification
became true:
\[ t_e = max(\{ t \mid \langle j, t \rangle \in J^T \}) \]

@@

The \emph{state} of a sliding-window $W^{S,r}$ at an instant
${i \in \mathbb{T}}$ is the \emph{set} of fact instances that are
associated in $S$ with any instant in the interval ${(i-r,i]}$, and is
denoted ${W^{S,r}_{i}}$.
The state of a sliding-window is defined as follows:
\[
W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
\]  
\end{definition}

@@

Axiom \ref{axiom:continuous datalog: window range leq 0} is trivially
true by reference to the interval in \refdef{continuous datalog: window},
and the definition of numeric intervals.  The state of a window of range
$0$ over stream $S$ at interval $i$ is composed of fact instances that
are each mapped by $S$ to any instant in the interval ${(i,i]}$.  This
interval is defined as the set of numbers ${\{ x \mid i < x \leqslant i \}}$,
which is the empty set for any $i$.

@@

{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
If two states of windows over the same in-order stream $S$ cover
non-overlapping time periods, then the set of fact instances entailed
by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
where ${z \geqslant x}$.
\[
\{ e \mid \langle R , W^{S,x}_i \rangle \models e \} \cap \{ e \mid \langle R , W^{S,y}_{i-z} \rangle \models e \} = \varnothing
\]
\end{axiom}}

@@

The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
semantically distinct portions: the \emph{persistent} portion is the
entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
program, which is guaranteed to be part of the entailment of every consistent
state due to the monotonicity of positive Datalog; the \emph{transient} portion
is the set $E^T_{i}$ of entailed fact instances whose justifications are at
least partially contained in $W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i$.
{\nobreak We define $E^T_i$ as:
\[
E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cap \left( W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i \right) \neq \varnothing \}
\]
where $J^T$ is a justification for the entailment of $x$ by
$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
  datalog: entailment disjointness}, $E^0$ is guaranteed to be
disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
WC-Datalog program $\Pi^{C,r}$.

>>>>

We define the concept of entailment in Datalog as follows. A fact $e$
is entailed by a program $\Pi$ iff each interpretation satisfying
$\Pi$ (satisfying every clause in $\Pi$) also satisfies $e$, written
$\Pi \models e$. For a program $\Pi$ and a set of entailed facts
$E$, $\Pi \models E$ iff $\Pi \models e$ for every $e \in E$. For
convenience, we refer to the set of all facts that are entailed by a
program (the intersection of all the Herbrand models of the program)
as the {\em entailment of a program}, written $cons(\Pi)$ and defined
as:
\begin{align*}
  cons(\Pi) &= \left\{ e \in \hat{H} \,\middle\vert\, \Pi \models e \right\} \\
  &= \bigcap_\mathcal{I} \models_\mathcal{I} \Pi
\end{align*}
\end{definition}

@@

For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
of fact instances and $F = \left\{ strip(f) \,\middle\vert\, f \in F^T \right\}$.

@@

The {\em entailment of the TA-Datalog program} $\Pi^T$ is therefore:
\[ cons^T(\Pi^T) = \left\{ \langle e, t_e \rangle \in \hat{H}^T \,\middle\vert\,
\Pi^T \models \langle e, t_e \rangle \right\} \]

@@

For an entailed fact instance $\langle e, t_e\rangle$ and a
justification of that fact instance $J^T$, the instant $t_e$ from which
$\langle e, t_e \rangle$ is recognised to be true is, intuitively, taken to be the
latest instant from which any of the fact instances in the justification
became true:
\[ t_e = max\left(\left\{ t \,\middle\vert\, \langle j, t \rangle \in J^T \right\}\right) \]

@@

The \emph{state} of a sliding-window $W^{S,r}$ at an instant
${i \in \mathbb{T}}$ is the \emph{set} of fact instances that are
associated in $S$ with any instant in the interval ${(i-r,i]}$, and is
denoted ${W^{S,r}_{i}}$.
The state of a sliding-window is defined as follows:
\[
W^{S,r}_i = \left\{ \langle f, t \rangle \,\middle\vert\, \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \right\}
\]  
\end{definition}

@@

Axiom \ref{axiom:continuous datalog: window range leq 0} is trivially
true by reference to the interval in \refdef{continuous datalog: window},
and the definition of numeric intervals.  The state of a window of range
$0$ over stream $S$ at interval $i$ is composed of fact instances that
are each mapped by $S$ to any instant in the interval ${(i,i]}$.  This
interval is defined as the set of numbers ${\left\{ x \,\middle\vert\, i < x \leqslant i \right\}}$,
which is the empty set for any $i$.

@@

{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
If two states of windows over the same in-order stream $S$ cover
non-overlapping time periods, then the set of fact instances entailed
by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
where ${z \geqslant x}$.
\[
\left\{ e \,\middle\vert\, \langle R , W^{S,x}_i \rangle \models e \right\} \cap \left\{ e \,\middle\vert\, \langle R , W^{S,y}_{i-z} \rangle \models e \right\} = \varnothing
\]
\end{axiom}}

@@

The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
semantically distinct portions: the \emph{persistent} portion is the
entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
program, which is guaranteed to be part of the entailment of every consistent
state due to the monotonicity of positive Datalog; the \emph{transient} portion
is the set $E^T_{i}$ of entailed fact instances whose justifications are at
least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i$.
{\nobreak We define $E^T_i$ as:
\[
E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_j,r_j}_i \neq \varnothing \right) \right\}
\]
where $J^T$ is a justification for the entailment of $e$ by
$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
  datalog: entailment disjointness}, $E^0$ is guaranteed to be
disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
WC-Datalog program $\Pi^{C,r}$.
